---
title: "ExploratoryData"
output: pdf_document
---

#1
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions 
## in the United States 
## from 1999 to 2008 as PNG in working directory

```{r}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```

## Q1
Have total emissions from PM2.5 decreased in the United States from 1999 to 2008?
Using the base plotting system, make a plot showing the total PM2.5 emission from all sources for each of the years 1999, 2002, 2005, and 2008.

```{r}
## organise the figures by year and scale them down
ds1 <- tapply(NEI$Emissions/1000, NEI$year, sum)
ds1.factor <- data.frame(year = names(ds1), 
  	Emissions = as.numeric(ds1))

## make the plot
barplot(ds1.factor$Emissions, names.arg = ds1.factor$year,
	ylab = "PM2.5 emissions (in thousand tons)", xlab = "Year",
	main = "Total PM2.5 Emissions in the United States from 1999 to 2008")

```
#### The total emissions of PM2.5 in the US from 1999 to 2008 have steadily decreased.

#2
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions
## in the Baltimore City, Maryland 
## from 1999 to 2008 as PNG in working directory

```{r, echo=FALSE}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```

## Q2
#### Have total emissions from PM2.5 decreased in the Baltimore City, Maryland (fips == "24510") 
#### from 1999 to 2008? Use the base plotting system to make a plot answering this question.
```{r}
## get the emissions figures
ds2 <- NEI[which(NEI$fips == "24510"), c("Emissions","year")]

## organise the figures by year and scale them down
ds2.t <- tapply(ds2$Emissions/1000, ds2$year, sum)
ds2.factor <- data.frame(year = names(ds2.t), 
  	Emissions = as.numeric(ds2.t))
```
```{r, echo=FALSE}
## make the plot
png(filename = "plot2.png",
    width = 720, height = 640)
par(mar=c(5.1, 4.1, 4.1, 2.1))
barplot(ds2.factor$Emissions, names.arg = ds2.factor$year,
	ylab = "PM2.5 emissions (in thousand tons)", xlab = "Year",
	main = "Total PM2.5 Emissions in the Baltimore City, Maryland from 1999 to 2008")
dev.off()
```
#### The total emissions of PM2.5 in the Baltimore City, Maryland from 1999 to 2008 showed a decreasing trend with vigorous fluctuation.

#3
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions 
## by source types in the United States 
## from 1999 to 2008 as PNG in working directory

```{r, echo=FALSE}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```

## Q3
#### Of the four types of sources indicated by the type (point, nonpoint, onroad, nonroad) variable, 
#### which of these four sources have seen decreases in emissions from 1999¡V2008 for Baltimore City? 
#### Which have seen increases in emissions from 1999¡V2008? 
```{r, echo=FALSE}
#### Use the ggplot2 plotting system to make a plot answer this question.

library(ggplot2)
```
```{r}
## get the data
ds3 <- NEI[which(NEI$fips == "24510"), c("Emissions","year","type")]

## get the total numbers
ds3.agg <- aggregate(Emissions ~ type + year, data = ds3, FUN = sum)
ds3.agg$Emissions<- as.factor(ds3.agg$Emissions)
```
```{r, echo=FALSE}
## make the plot
png(filename = "plot3.png",
    width = 1024, height = 720)
ggplot(ds3.agg,aes(year,Emissions,fill=type)) +
  scale_x_continuous(breaks=c(1999, 2002, 2005, 2008)) +
	geom_bar(stat="identity",position='dodge')+
	xlab("Year") +
	ylab("Total PM2.5 emissions in tons") +
	ggtitle("Total PM2.5 Emissions in the Baltimore City, Maryland from 1999 to 2008")
dev.off()
```

## From 1999¡V2008 for Baltimore City, the 4 types of PM2.5 emissions showed a descreasing tendency.
## PM2.5 emissions from NONPOINT sources slightly decreased and remained the greatest emissions over the period.
## PM2.5 emissions from ON-ROAD sources steadily decreased.
## PM2.5 emissions from  POINT sources increased with fluctuations.
## PM2.5 emissions from NON-ROAD sources decreased with fluctuations and became the least emissions sources in 2008.

#4
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions
## from coal combustion-related sources
## in the United States from 1999 to 2008 as PNG in working directory
```{r, echo=FALSE}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```
## Q4
#### Across the United States, how have emissions from coal combustion-related sources changed from 1999¡V2008?
```{r}
## get the SCC code for coal consumption related records
coalcode <- SCC[grep(".*[cC]oal.*", SCC$EI.Sector), c("SCC")]

## get the data according to the SCC codes
ds4 <- NEI[NEI$SCC %in% coalcode, c("Emissions","year")]

## organise the figures by year and scale them down
ds4.t <- tapply(ds4$Emissions/1000, ds4$year, sum)
ds4.factor <- data.frame(year = names(ds4.t), 
  	Emissions = as.numeric(ds4.t))
```
```{r, echo=FALSE}
## make the plot
png(filename = "plot4.png",
    width = 1040, height = 720)
par(mar=c(5.1, 4.1, 4.1, 2.1))
barplot(ds4.factor$Emissions, names.arg = ds4.factor$year,
	ylab = "PM2.5 emissions (in thousand tons)", xlab = "Year",
	main = "Total PM2.5 Emissions from Coal Combustion-related Sources in the United States from 1999 to 2008")
dev.off()
```

## The PM2.5 emissions showed a decreasing trend throughout the period 
## and a remarkable drop from 2005 to 2008, which was approximately one-third of the emissions in 1999.

#5
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions
## from motor vehicle sources in Baltimore City
## from 1999 to 2008 as PNG in working directory.
```{r, echo=FALSE}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```
## Q5
#### How have emissions from motor vehicle sources changed from 1999¡V2008 in Baltimore City?
```{r}
## get the SCC code for motor vehicle sources records
motorcode <- SCC[grep(".*[mM]otor.*", SCC$EI.Sector), c("SCC")]
vehiclecode <- SCC[grep(".*[vV]ehicle.*", SCC$EI.Sector), c("SCC")]

## get the data according to the SCC codes
ds5 <- NEI[which(NEI$fips == "24510"), c("Emissions","year","SCC")]
ds5 <- ds5[which(ds5$SCC %in% motorcode | ds5$SCC %in% vehiclecode),]

## organise the figures by year
ds5.t <- tapply(ds5$Emissions, ds5$year, sum)
ds5.factor <- data.frame(year = names(ds5.t), 
  	Emissions = as.numeric(ds5.t))
```
```{r, echo=FALSE}
## make the plot
png(filename = "plot5.png",
    width = 1040, height = 720)
par(mar=c(5.1, 4.1, 4.1, 2.1))
barplot(ds5.factor$Emissions, names.arg = ds5.factor$year,
	ylab = "PM2.5 emissions (in tons)", xlab = "Year",
	main = "Total PM2.5 Emissions from Motor Vehicle Sources in Baltimore City, Maryland from 1999 to 2008")
dev.off()
```

## The PM2.5 emissions showed a promising decreasing trend.
## The figure dropped more than halved from 1999 to 2002
## and then moderately decreased from 2002 to 2008.

#6
## This program assumes the required files are downloaded in working directory
## and the libraries are present.
## It creates a bar chart showing total PM2.5 emissions
## from motor vehicle sources in Baltimore City and Los Angeles County
## from 1999 to 2008 as PNG in working directory.
```{r, echo=FALSE}
## Read the data
NEI <- readRDS("summarySCC_PM25.rds")
SCC <- readRDS("Source_Classification_Code.rds")
```
## Q6
#### Compare emissions from motor vehicle sources in Baltimore City 
#### with emissions from motor vehicle sources in Los Angeles County, California (fips == "06037"). 
#### Which city has seen greater changes over time in motor vehicle emissions?
```{r}
## get the SCC code for motor vehicle sources records
motorcode <- SCC[grep(".*[mM]otor.*", SCC$EI.Sector), c("SCC")]
vehiclecode <- SCC[grep(".*[vV]ehicle.*", SCC$EI.Sector), c("SCC")]

## get the data according to the fipds and SCC codes
ds6 <- NEI[which(NEI$fips == "24510" | NEI$fips == "06037"), c("Emissions","year","SCC","fips")]
ds6 <- ds6[which(ds6$SCC %in% motorcode | ds6$SCC %in% vehiclecode),]
ds6[which(ds6$fips == "24510"), c("fips")] <- "Baltimore"
ds6[which(ds6$fips == "06037"), c("fips")] <- "Los Angeles County"

library(ggplot2)

## get the total numbers
ds6.agg <- aggregate(Emissions ~ fips + year, data = ds6, FUN = sum)
ds6.agg$Emissions<- as.factor(ds6.agg$Emissions)
```
```{r, echo=FALSE}
## make the plot
png(filename = "plot6.png",
    width = 1024, height = 720)
par(mar=c(5.1, 4.1, 4.1, 2.1))
ggplot(ds6.agg,aes(year,Emissions,fill=fips)) +
  scale_x_continuous(breaks=c(1999, 2002, 2005, 2008)) +
	geom_bar(stat="identity",position='dodge')+
	xlab("Year") +
	ylab("Total PM2.5 emissions in tons") +
	ggtitle("Total PM2.5 Emissions from Motor Vehicle Sources in Baltimore City, Maryland and Los Angeles County from 1999 to 2008") + 
	theme(plot.title = element_text(hjust = 0.5))
dev.off()
```
## The PM2.5 emissions in Baltimore City steadily decreased over the 10 years.
## Meanwhile the emissions in L. A. County rose significantly from 1999 to 2005 
## although a slight decline followed until 2008.
## L. A. County always emitted more than Baltimore City did by 10-50 times during the period.
## The difference was enlarging by the drop in Baltimore and the increase in L. A..
